---
title: "Results and Analysis for gcamland paper"
output: bookdown::word_document2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
library(cowplot)
```

```{r global-assumptions, echo=FALSE}
CROPS_TO_PLOT <- c("Corn", "Wheat", "OilCrop", "OtherGrain")
ALL_CROPS_NF <- c("Corn", "FiberCrop", "MiscCrop", "OilCrop", "OtherGrain",  "Rice", "Root_Tuber", "SugarCrop",  "Wheat")
CROP_GROUP1 <- "Corn, OilCrop"
CROP_GROUP2 <- "Wheat, OtherGrain"
CROP_GROUP3 <- "All Other Crops"
SCENARIOS_TO_PLOT <- c("Adaptive", "Linear", "Perfect", "Hybrid Linear Adaptive") # Could also include `GCAM_Default`
theme_set(theme_bw())
```

```{r global-data, echo=FALSE}
# Read data
best_models <- read.csv("./2-process/best_models.csv")
best_output <- read.csv("./2-process/best_output.csv")
all_outputs <- readRDS("./1-data/output_summarized.rds")
obs_data <- read.csv("./2-process/obs_data.csv")

# Rename expectations
best_models %>%
  mutate(expectation.type = sub("LaggedCurr", "Hybrid Perfect Adaptive", expectation.type),
         expectation.type = sub("Lagged", "Adaptive", expectation.type),
         expectation.type = sub("Mixed", "Hybrid Linear Adaptive", expectation.type),
         expectation.type = sub("HybridLinearAdaptive", "Hybrid Linear Adaptive", expectation.type)) ->
  best_models
best_output %>%
  mutate(scenario = sub("Best_", "", scenario),
         scenario = sub("LaggedCurr", "Hybrid Perfect Adaptive", scenario),
         scenario = sub("Lagged", "Adaptive", scenario),
         scenario = sub("Mixed", "Hybrid Linear Adaptive", scenario),
         scenario = sub("HybridLinearAdaptive", "Hybrid Linear Adaptive", scenario)) ->
  best_output

# Filter for the expectations included in this paper
# Note: Hybrid Perfect Expectations may make sense in the future as a imperfect forecast, but 
#       it has little justification for historical analysis
best_models %>%
  filter(expectation.type %in% SCENARIOS_TO_PLOT) ->
  best_models
best_output %>%
  filter(scenario %in% SCENARIOS_TO_PLOT) ->
  best_output
```

```{r colors, echo=FALSE}
CROP_COLORS <- c('#a6cee3','#1f78b4','#b2df8a','#33a02c','#fb9a99','#e31a1c','#fdbf6f','#ff7f00','#cab2d6','#6a3d9a','#b15928')
ET_COLORS <- c('Hybrid Perfect Adaptive' = '#7fc97f', 'Adaptive' = '#f0027f', 'Linear' = '#beaed4', 'Perfect' = '#fdc086','Hybrid Linear Adaptive' = '#386cb0', 'Obs' = 'black')
ASSUME_COLORS <- c('Default' = '#f0027f', 'With Subsidy' = '#7fc97f', 'Same Parameters' = '#beaed4')
OBJ_COLORS <- c('Default (nrms)' = '#f0027f', 'rms' = '#7fc97f', 'kge' = '#beaed4', 'bias' = '#fdc086')
LTY_COLORS <- c('Default (All Crops)' = '#f0027f', 'Corn' = '#7fc97f', 'OilCrop' = '#beaed4', 'Wheat' = '#fdc086', 'OtherGrain' = '#386cb0')
BY_COLORS <- c('Default (1990)' = '#f0027f', '1975' = '#7fc97f', '2005' = '#beaed4')
```

#### 1. Introduction

```{r hist-land-crop, echo=FALSE, fig.width=10, fig.height=6, fig.cap="Harvested area by crop for major commodities in the United States (1975-2015). Source: USDA"}
read.csv("./1-data/all_usda.csv") %>%
  filter(Year %in% 1975:2015,
         Type == "Harvested") ->
  usda

# =========
# Plot harvested area time series
x.label <- "Year"
y.label <- expression(paste("Area (thous ", km^2, ")"))
p <- ggplot() + geom_area(data=usda, aes(Year, Value, fill=GCAM_commodity))
p <- p + xlab(x.label) + ylab(y.label)
p <- p + scale_fill_manual(name="Commodity", values=CROP_COLORS)
p <- p + theme( legend.text = element_text(size = 12),
                legend.title = element_text(size = 14),
                strip.text = element_text(size = 14, face="bold"), 
                axis.text = element_text(size = 12), 
                axis.title = element_text(size = 14) )
print(p)
ggsave("./4-figures/Figure1.png", dpi=500, height=6, width=10)
```

## 4. Results

### 4.1. Parameter Sets that Minimize NRMS in gcamland

```{r param-n-main, echo=FALSE}
best_models %>%
  filter(calibration.year == 1990,
         comparison.years == "1990-2015",
         comparison.landTypes == "All Crops",
         assumption == "Default",
         objfun == "nrms") ->
  nrms
```

#### Values for paragraph on NRMS

Smallest NRMS is `r round(min(nrms$landTypeMeanObjFunVal), 3)` with `r nrms[which.min(nrms$landTypeMeanObjFunVal), "expectation.type"]` expectations.

Largest NRMS is `r round(max(nrms$landTypeMeanObjFunVal), 3)` with `r nrms[which.max(nrms$landTypeMeanObjFunVal), "expectation.type"]` expectations. 

#### Parameters in Default Model

```{r param-l-main, echo=FALSE, warning=FALSE, message=FALSE, fig.width=10, fig.height=6, fig.cap="Parameters that minimize NRMSE by Expectation Type"}
nrms %>%
  select(expectation.type, logit.agforest, logit.afnonpast, logit.crop, share.old1, share.old2, share.old3, linear.years1, linear.years2, linear.years3) %>%
  gather(parameter, value, -expectation.type) %>%
  separate(parameter, into=c("param_type", "name")) %>%
  mutate(param_type = if_else(param_type == "logit", "Logit Exponent", param_type),
         param_type = if_else(param_type == "share", "Share of Past Information", param_type),
         param_type = if_else(param_type == "linear", "Number of Years in Linear Forecast", param_type),
         name = if_else(grepl("1", name), CROP_GROUP1, name),
         name = if_else(grepl("2", name), CROP_GROUP2, name),
         name = if_else(grepl("3", name), CROP_GROUP3, name),
         name = if_else(name == "crop", "Cropland Nest", name),
         name = if_else(name == "agforest", "Dynamic Land Nest", name),
         name = if_else(name == "afnonpast", "Ag, Forest, and Other Nest", name)) ->
  parameters

parameters$name <- factor(parameters$name,levels=c( "Dynamic Land Nest", "Ag, Forest, and Other Nest", "Cropland Nest", CROP_GROUP1, CROP_GROUP2,  CROP_GROUP3))
p <- ggplot() + geom_bar(data=parameters, aes(name, value, fill=expectation.type), stat="identity", position = "dodge")
p <- p + facet_wrap(~param_type, scales="free", labeller = label_wrap_gen())
p <- p + scale_fill_manual(name="Expectation Type", values=ET_COLORS)
p <- p + theme(axis.text.x=element_text(angle=90,hjust=1))
p <- p + xlab("") + ylab("Value (various units)")
p <- p + theme( legend.text = element_text(size = 12),
                legend.title = element_text(size = 14),
                strip.text = element_text(size = 14, face="bold"), 
                axis.text = element_text(size = 12), 
                axis.title = element_text(size = 14) )
print(p)
ggsave("./4-figures/Figure3.png", dpi=500, height=6, width=10)
```


### 4.2. Comparing Modeled Land Area to Observations

```{r crop-nrms, echo=FALSE}
read.csv("./2-process/snyder_comparison_table.csv") %>%
  filter(grepl("ThisPaper", Type)) ->
  crop.nrms
```


#### Values for paragraph on crop-specific NRMS

The NRMS for Corn and OilCrop are larger (`r round(crop.nrms[crop.nrms$crop == "Corn" & crop.nrms$Type == "ThisPaper_AllCrops", "value"],2)` and `r round(crop.nrms[crop.nrms$crop == "OilCrop" & crop.nrms$Type == "ThisPaper_AllCrops", "value"],2)`, respectively) than the NRMS for other Wheat and OtherGrain (`r round(crop.nrms[crop.nrms$crop == "Wheat" & crop.nrms$Type == "ThisPaper_AllCrops", "value"], 2)` and `r round(crop.nrms[crop.nrms$crop == "OtherGrain" & crop.nrms$Type == "ThisPaper_AllCrops", "value"], 2)`, respectively).

#### Cropland area over time in default model

```{r crop-ts-main, echo=FALSE, message=FALSE, fig.width=10, fig.height=6, fig.cap="Harvested cropland area (total and by crop) over time by expectation type. Black line is observations (FAO). Colored lines are gcamland results for the models that minimize NRMSE. The expectation type with the minimum NRMSE (Adaptive) is shown with a thicker line. Gray area is the range of all gcamland simulations. Note that fodder crops are included in gcamland but are excluded from total cropland area in this figure due to data limitations. "}
# Calculate total cropland from the best default models
best_output %>%
  filter(calibration.year == 1990,
         comparison.years == "1990-2015",
         comparison.landTypes == "All Crops",
         assumption == "Default",
         objfun == "nrms") %>%
  filter(GCAM_commodity %in% ALL_CROPS_NF) %>%
  filter(scenario %in% SCENARIOS_TO_PLOT) %>%
  group_by(scenario, year) %>%
  summarize(harvested.land = sum(harvested.land)) %>%
  ungroup() %>%
  mutate(GCAM_commodity = "Harvested crop area") ->
  gcam

best.type <- nrms[which.min(nrms$landTypeMeanObjFunVal), "expectation.type"]
gcam %>%
  filter(scenario == best.type) ->
  gcam_best

# Filter all outputs for the right information
all_outputs %>%
  mutate(name = if_else(name == "crops_nonfodder", "Harvested crop area", name)) %>%
  rename(GCAM_commodity = name) %>%
  filter(GCAM_commodity %in% c("Harvested crop area")) ->
  gcam_ribbon

# Filter obs data for the right information
obs_data %>%
  filter(region == "USA", 
         GCAM_commodity %in% c("Cropland_Harvested"),
         year %in% 1975:2015) %>%
  mutate(GCAM_commodity = "Harvested crop area") ->
  obs

y.label <- expression(paste("Area (thous ", km^2, ")"))
p1 <- ggplot() + geom_ribbon(data=gcam_ribbon, aes(x=year, ymin=min_land, ymax=max_land), color="gray", alpha=0.2)
p1 <- p1 + geom_line(data=obs, aes(year, area), color="black", size=2)
p1 <- p1 + geom_line(data=gcam_best, aes(year, harvested.land, color=scenario), size=2)
p1 <- p1 + geom_line(data=gcam, aes(year, harvested.land, color=scenario), size=1)
p1 <- p1 + xlab("Year") + ylab(y.label) + scale_color_manual(name="Expectation Type", values=ET_COLORS)
p1 <- p1 + facet_wrap(~GCAM_commodity, scales="free_y")
p1 <- p1 + scale_x_continuous(breaks=seq(1975, 2010, 15)) 
p1 <- p1 + coord_cartesian(ylim = c(0, 2000)) 
p1 <- p1 + theme(legend.position = "none",
                strip.text = element_text(size = 14, face="bold"), 
                axis.text = element_text(size = 12), 
                axis.title = element_text(size = 14) )

# Repeat for specific crops
all_outputs %>%
  rename(GCAM_commodity = name) %>%
  filter(GCAM_commodity %in% CROPS_TO_PLOT) ->
  gcam_ribbon
best_output %>%
  filter(calibration.year == 1990,
         comparison.years == "1990-2015",
         comparison.landTypes == "All Crops",
         assumption == "Default",
         objfun == "nrms") %>%
  filter(GCAM_commodity %in% CROPS_TO_PLOT) %>%
  filter(scenario %in% SCENARIOS_TO_PLOT) ->
  gcam
gcam %>%
  filter(scenario == best.type) ->
  gcam_best
obs_data %>%
  filter(region == "USA", 
         GCAM_commodity %in% CROPS_TO_PLOT,
         year %in% 1975:2015) ->
  obs

y.label <- expression(paste("Area (thous ", km^2, ")"))
p2 <- ggplot() + geom_ribbon(data=gcam_ribbon, aes(x=year, ymin=min_land, ymax=max_land), color="gray", alpha=0.2)
p2 <- p2 + geom_line(data=obs, aes(year, area), color="black", size=2)
p2 <- p2 + geom_line(data=gcam_best, aes(year, harvested.land, color=scenario), size=2)
p2 <- p2 + geom_line(data=gcam, aes(year, harvested.land, color=scenario), size=1)
p2 <- p2 + xlab("Year") + ylab("") + scale_color_manual(name="Expectation Type", values=ET_COLORS)
p2 <- p2 + facet_wrap(~GCAM_commodity, scales="free_y")
p2 <- p2 + scale_x_continuous(breaks=seq(1975, 2015, 15)) 
p2 <- p2 + coord_cartesian(ylim = c(0, 500)) 
p2 <- p2 + theme( legend.text = element_text(size = 12),
                legend.title = element_text(size = 14),
                strip.text = element_text(size = 14, face="bold"), 
                axis.text = element_text(size = 12), 
                axis.title = element_text(size = 14) )
 
plot_grid(p1, p2, ncol = 2, rel_widths = c(1/3, 2/3))
ggsave("./4-figures/Figure4.png", dpi=500, height=6, width=10)
```

## 5. Sensitivity Analysis

### 5.1. Sensitivity to Model Assumptions


```{r param-assume-sens, warning=FALSE, message=FALSE, echo=FALSE}
best_models %>%
  filter(calibration.year == 1990,
         comparison.years == "1990-2015",
         comparison.landTypes == "All Crops",
         objfun == "nrms",
         assumption != "5 year timestep") ->
  TEMP
TEMP %>% 
  group_by(assumption, objfun, calibration.year, comparison.years, comparison.landTypes) %>%
  summarize(landTypeMeanObjFunVal = min(landTypeMeanObjFunVal)) %>%
  ungroup() %>%
  left_join(select(TEMP, assumption, expectation.type, landTypeMeanObjFunVal), by=c("assumption", "landTypeMeanObjFunVal"))->
  assumptions
```

#### Values for paragraph on assumption sensitivity

Expectation type in default model = `r assumptions[assumptions$assumption == "Default", "expectation.type"]`

Expectation type with subsidies = `r assumptions[assumptions$assumption == "With Subsidy", "expectation.type"]`

Expectation type with same param = `r assumptions[assumptions$assumption == "Same Parameters", "expectation.type"]` 

NRMS in default = `r round(assumptions[assumptions$assumption == "Default", "landTypeMeanObjFunVal"], 3)` 

NRMS with subsidies =  `r round(assumptions[assumptions$assumption == "With Subsidy", "landTypeMeanObjFunVal"], 3)` 

NRMS with same parameters `r round(assumptions[assumptions$assumption == "Same Parameters", "landTypeMeanObjFunVal"], 3)`). 

#### Cropland with different assumptions

```{r crop-assume-sens, echo=FALSE, fig.width=10, fig.height=6, fig.cap="Harvested area by crop under different model assumptions. Black line is observations (FAO). Colored lines are gcamland results for the models that minimize NRMSE."}
assumptions %>%
  select(assumption, expectation.type) %>%
  rename(best.type = expectation.type) ->
  best_assumptions

best_output %>%
  filter(calibration.year == 1990,
         comparison.years == "1990-2015",
         comparison.landTypes == "All Crops",
         objfun == "nrms",
         assumption != "5 year timestep") %>%
  mutate(scenario = sub("Best_", "", scenario)) %>%
  left_join(best_assumptions, by=c("assumption")) %>%
  filter(scenario == best.type) %>%
  filter(GCAM_commodity %in% CROPS_TO_PLOT) ->
  gcam
gcam %>%
  filter(assumption == "Default") ->
  gcam_default

obs_data %>%
  filter(region == "USA", 
         GCAM_commodity %in% CROPS_TO_PLOT,
         year %in% 1975:2015) ->
  obs

y.label <- expression(paste("Area (thous ", km^2, ")"))
p <- ggplot() + geom_line(data=obs, aes(year, area), color="black", size=1.5)
p <- p + geom_line(data=gcam_default, aes(year, harvested.land, color=assumption, lty=scenario), size=1.5)
p <- p + geom_line(data=gcam, aes(year, harvested.land, color=assumption, lty=scenario), size=1)
p <- p + xlab("Year") + ylab(y.label) + scale_color_manual(name="", values=ASSUME_COLORS)
p <- p + facet_wrap(~GCAM_commodity, scales="free_y")
p <- p + scale_x_continuous(breaks=seq(1975, 2015, 15)) 
p <- p + coord_cartesian(ylim = c(0, 500)) 
p <- p + theme( legend.text = element_text(size = 12),
                legend.title = element_text(size = 14),
                strip.text = element_text(size = 14, face="bold"), 
                axis.text = element_text(size = 12), 
                axis.title = element_text(size = 14) )
print(p)
ggsave("./4-figures/Figure5.png", dpi=500, height=6, width=10)
```

### 5.2. Sensitivity to the Objective Function


#### 5.2.1. Optimizing for different objective functions


```{r param-obj-sens, echo=FALSE, message=FALSE, warning=FALSE}
best_models %>%
  filter(calibration.year == 1990,
         comparison.years == "1990-2015",
         comparison.landTypes == "All Crops",
         assumption == "Default") ->
  TEMP
TEMP %>% 
  group_by(assumption, objfun, calibration.year, comparison.years, comparison.landTypes) %>%
  summarize(landTypeMeanObjFunVal = min(landTypeMeanObjFunVal)) %>%
  ungroup() %>%
  left_join(select(TEMP, objfun, expectation.type, landTypeMeanObjFunVal), by=c("objfun", "landTypeMeanObjFunVal"))->
  objective
```

#### Cropland area with different objective functions
```{r crop-obj-sens, echo=FALSE, fig.width=10, fig.height=6, fig.cap="Harvested area by crop when optimizing for different objective functions. Colors indicate objective function. Line type indicates the expectation type that minimizes that objective function. Only the objective function minimizing expectation type is shown. Note that NRMSE and RMSE result in identical parameter sets in the default model and thus have identical land allocation in this figure."}
objective %>%
  select(objfun, expectation.type) %>%
  rename(best.type = expectation.type) ->
  best_assumptions

best_output %>%
  filter(calibration.year == 1990,
         comparison.years == "1990-2015",
         comparison.landTypes == "All Crops",
         assumption == "Default") %>%
  mutate(scenario = sub("Best_", "", scenario)) %>%
  left_join(best_assumptions, by=c("objfun")) %>%
  filter(scenario == best.type,
         GCAM_commodity %in% CROPS_TO_PLOT) %>%
  mutate(objfun = if_else(objfun == "nrms", "Default (nrms)", objfun)) ->
  gcam
gcam %>%
  filter(objfun == "Default (nrms)") ->
  gcam_default

obs_data %>%
  filter(region == "USA", 
         GCAM_commodity %in% CROPS_TO_PLOT,
         year %in% 1975:2015) ->
  obs

y.label <- expression(paste("Area (thous ", km^2, ")"))
p <- ggplot() + geom_line(data=obs, aes(year, area), lty="solid", color="black", size=1.5)
p <- p + geom_line(data=gcam_default, aes(year, harvested.land, color=objfun, lty=scenario), size=1.5)
p <- p + geom_line(data=gcam, aes(year, harvested.land, color=objfun, lty=scenario), size=1)
p <- p + xlab("Year") + ylab(y.label) + scale_color_manual(name="", values=OBJ_COLORS)
p <- p + scale_linetype_manual(name="", values=c("solid", "dashed", "dotted"))
p <- p + facet_wrap(~GCAM_commodity, scales="free_y")
p <- p + scale_x_continuous(breaks=seq(1975, 2015, 15))  
p <- p + coord_cartesian(ylim = c(0, 600)) 
p <- p + guides(color = guide_legend(override.aes = list(size = 1)),
         linetype = guide_legend(override.aes = list(size=1)),
         size = FALSE)
p <- p + theme( legend.text = element_text(size = 12),
                legend.title = element_text(size = 14),
                legend.key.size = unit(1, 'cm'),
                strip.text = element_text(size = 14, face="bold"), 
                axis.text = element_text(size = 12), 
                axis.title = element_text(size = 14) )
print(p)
ggsave("./4-figures/Figure6.png", dpi=500, height=6, width=10)
```

#### 5.2.2. Optimizing for different land types

#### Values for paragraph on NRMS

NRMS for Corn is reduced from `r round(crop.nrms[crop.nrms$crop == "Corn" & crop.nrms$Type == "ThisPaper_AllCrops", "value"],2)` to `r round(crop.nrms[crop.nrms$crop == "Corn" & crop.nrms$Type == "ThisPaper_SingleCrop", "value"],2)` when we go from minimizing NRMS across all crops to minimizing NRMS for Corn only. 

NRMS for OilCrop is reduced from `r round(crop.nrms[crop.nrms$crop == "OilCrop" &crop.nrms$Type == "ThisPaper_AllCrops", "value"],2)` to `r round(crop.nrms[crop.nrms$crop == "OilCrop" &crop.nrms$Type == "ThisPaper_SingleCrop", "value"],2)` when we go from minimizing NRMS across all crops to minimizing NRMS for OilCrop only. 

Optimizing for a single crop has less effect on the NRMS for Wheat and OtherGrain (from `r round(crop.nrms[crop.nrms$crop == "Wheat" &crop.nrms$Type == "ThisPaper_AllCrops", "value"],2)` to `r round(crop.nrms[crop.nrms$crop == "Wheat" &crop.nrms$Type == "ThisPaper_SingleCrop", "value"],2)` for Wheat, and from `r round(crop.nrms[crop.nrms$crop == "OtherGrain" &crop.nrms$Type == "ThisPaper_AllCrops", "value"],2)` to `r round(crop.nrms[crop.nrms$crop == "OtherGrain" &crop.nrms$Type == "ThisPaper_SingleCrop", "value"],2)` for OtherGrain), most likely because the performance in the all crops is sufficiently good.

```{r param-lty-sens, echo=FALSE, message=FALSE, warning=FALSE}
best_models %>%
  filter(calibration.year == 1990,
         comparison.years == "1990-2015",
         assumption == "Default",
         objfun == "nrms",
         !(comparison.landTypes %in% c("Crops, excluding PalmFruit", "All dynamic land types"))) ->
  TEMP
TEMP %>% 
  group_by(assumption, objfun, calibration.year, comparison.years, comparison.landTypes) %>%
  summarize(landTypeMeanObjFunVal = min(landTypeMeanObjFunVal)) %>%
  ungroup() %>%
  left_join(select(TEMP, comparison.landTypes, expectation.type, landTypeMeanObjFunVal), by=c("comparison.landTypes", "landTypeMeanObjFunVal"))->
  lty_param
```

Including land cover types increases NRMSE from `r round(lty_param[lty_param$comparison.landTypes == "All Crops" & lty_param$expectation.type == "Adaptive", "landTypeMeanObjFunVal"],2)` to `r round(best_models[best_models$comparison.landTypes == "All dynamic land types" & best_models$expectation.type == "Adaptive", "landTypeMeanObjFunVal"],2)`. However, parameters do not change. For example, the logit exponent over the dynamic land nest is `r best_models[best_models$comparison.landTypes == "All Crops" & best_models$expectation.type == "Adaptive" & best_models$calibration.year == 1990 & best_models$objfun == "nrms" & best_models$assumption == "Default", "logit.agforest"]` in the default model and `r best_models[best_models$comparison.landTypes == "All dynamic land types" & best_models$expectation.type == "Adaptive" & best_models$calibration.year == 1990 & best_models$objfun == "nrms", "logit.agforest"]` when all dynamic land types are included in the NRMSE calculation.

#### Cropland area when optimizing for different land types

```{r crop-lty-sens, echo=FALSE, fig.width=10, fig.height=6, fig.cap="Harvested area by crop when optimizing for different land types. Colors indicate crops included in the objective function. Line type indicates the expectation type that minimizes NRMSE for that set of crops. Only the NRMSE minimizing expectation type for each set of crops is shown."}
lty_param %>%
  select(comparison.landTypes, expectation.type) %>%
  rename(best.type = expectation.type) ->
  best_assumptions

best_output %>%
  filter(calibration.year == 1990,
         comparison.years == "1990-2015",
         assumption == "Default",
         objfun == "nrms",
         !(comparison.landTypes %in% c("Crops, excluding PalmFruit", "All dynamic land types"))) %>%
  mutate(scenario = sub("Best_", "", scenario)) %>%
  left_join(best_assumptions, by=c("comparison.landTypes")) %>%
  filter(GCAM_commodity %in% CROPS_TO_PLOT,
         scenario == best.type) %>%
  mutate(comparison.landTypes = if_else(comparison.landTypes == "All Crops", "Default (All Crops)", comparison.landTypes) )->
  gcam
gcam %>%
  filter(comparison.landTypes == "Default (All Crops)") ->
  gcam_default
write.csv(gcam, "~/Documents/gcamland.csv", row.names=FALSE)
obs_data %>%
  filter(region == "USA", 
         GCAM_commodity %in% CROPS_TO_PLOT,
         year %in% 1975:2015) ->
  obs

y.label <- expression(paste("Area (thous ", km^2, ")"))
p <- ggplot() + geom_line(data=gcam_default, aes(year, harvested.land, color=comparison.landTypes, lty=scenario), size=1.5)
p <- p + geom_line(data=obs, aes(year, area), color="black", size=1.5)
p <- p + geom_line(data=gcam, aes(year, harvested.land, color=comparison.landTypes, lty=scenario), size=1)
p <- p + xlab("Year") + ylab(y.label) + scale_color_manual(name="", values=LTY_COLORS)
p <- p + scale_linetype_manual(name="", values=c("solid", "dashed", "dotted"))
p <- p + facet_wrap(~GCAM_commodity)
p <- p + scale_x_continuous(breaks=seq(1975, 2015, 15))  
p <- p + coord_cartesian(ylim = c(0, 650)) 
p <- p + guides(color = guide_legend(override.aes = list(size = 1)),
         linetype = guide_legend(override.aes = list(size=1)),
         size = FALSE)
p <- p + theme( legend.text = element_text(size = 12),
                legend.title = element_text(size = 14),
                legend.key.size = unit(1, 'cm'),
                strip.text = element_text(size = 14, face="bold"), 
                axis.text = element_text(size = 12), 
                axis.title = element_text(size = 14) )
print(p)
ggsave("./4-figures/Figure7.png", dpi=500, height=6, width=10)
```

### 5.3. Different calibration years


```{r param-time-sens, echo=FALSE, warning=FALSE, message=FALSE}
best_models %>%
  filter(comparison.landTypes == "All Crops",
         comparison.years == paste(calibration.year, "-2015", sep=""),
         assumption == "Default",
         objfun == "nrms") ->
  TEMP
TEMP %>% 
  group_by(assumption, objfun, calibration.year, comparison.years, comparison.landTypes) %>%
  summarize(landTypeMeanObjFunVal = min(landTypeMeanObjFunVal)) %>%
  ungroup() %>%
  left_join(select(TEMP, calibration.year, expectation.type, landTypeMeanObjFunVal), by=c("calibration.year", "landTypeMeanObjFunVal"))->
  calibYear
```

#### Cropland area with different calibration years

```{r crop-time-sens, echo=FALSE, fig.width=10, fig.height=6, fig.cap="Harvested area by crop when using different calibration years. Colors indicate calibration year. Line type indicates the expectation type that minimizes that NRMSE for that calibration year. Only the NRMSE minimizing expectation type for each calibration year is shown."}
calibYear %>%
  select(calibration.year, expectation.type) %>%
  rename(best.type = expectation.type) ->
  best_assumptions

best_output %>%
  filter(comparison.landTypes == "All Crops",
         comparison.years == paste(calibration.year, "-2015", sep=""),
         assumption == "Default",
         objfun == "nrms") %>%
  mutate(scenario = sub("Best_", "", scenario)) %>%
  left_join(best_assumptions, by=c("calibration.year")) %>%
  filter(GCAM_commodity %in% CROPS_TO_PLOT,
         scenario == best.type) %>%
  mutate(calibration.year = if_else(calibration.year == 1990, "Default (1990)", as.character(calibration.year))) ->
  gcam
gcam %>%
  filter(calibration.year == "Default (1990)") ->
  gcam_default

obs_data %>%
  filter(region == "USA", 
         GCAM_commodity %in% CROPS_TO_PLOT,
         year %in% 1975:2015) ->
  obs

y.label <- expression(paste("Area (thous ", km^2, ")"))
p <- ggplot() + geom_line(data=obs, aes(year, area), color="black", size=1.5)
p <- p + geom_line(data=gcam_default, aes(year, harvested.land, color=as.factor(calibration.year), lty=scenario), size=1.5)
p <- p + geom_line(data=gcam, aes(year, harvested.land, color=as.factor(calibration.year), lty=scenario), size=1)
p <- p + xlab("Year") + ylab(y.label) + scale_color_manual(name="", values=BY_COLORS)
p <- p + facet_wrap(~GCAM_commodity)
p <-  p + scale_x_continuous(breaks=seq(1975, 2015, 15)) 
p <- p + coord_cartesian(ylim = c(0, 500)) 
p <- p + scale_linetype_manual(name="", values=c("solid", "dashed", "dotted"))
p <- p + guides(color = guide_legend(override.aes = list(size = 1)),
         linetype = guide_legend(override.aes = list(size=1)),
         size = FALSE)
p <- p + theme( legend.text = element_text(size = 12),
                legend.title = element_text(size = 14),
                legend.key.size = unit(1, 'cm'),
                strip.text = element_text(size = 14, face="bold"), 
                axis.text = element_text(size = 12), 
                axis.title = element_text(size = 14) )
print(p)
ggsave("./4-figures/Figure8.png", dpi=500, height=6, width=10)
```



